---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r unzipping and read in}
unzip('activity.zip', list = FALSE, overwrite = TRUE)
activity.raw <- read.csv('activity.csv')
```

Since the Date Variable is actually imported as a factor we had to convert it 
to date.

```{r convert date}
activity.raw$date <- as.Date(activity.raw$date, "%Y-%m-%d")
```




## What is mean total number of steps taken per day?
First of to answer this question we have to omit the NA-Values, after that we can calculate the sum of steps per day.
Also we calculate the median and mean of the steps per day, because it is asked
in the second question.

```{r sum up the steps per day}

# Removing NAs
activity.wona <- activity.raw[complete.cases(activity.raw), ]

# summarize the data
library(dplyr)
activity.grpd <- group_by(activity.wona, date)
activity.daily <- summarise(activity.grpd,
  sum_of_steps = sum(steps, na.rm = TRUE),
  mean = mean(steps, na.rm = TRUE),
  median = median(steps, na.rm = TRUE))
```

After we got that data we need, we plot it in a historgram

```{r histogram}
hist(activity.daily$sum_of_steps, 
     main="Histogram of daily sum of steps",
     xlab="daily sum of steps")
```

Now we can answer the second question. For this we´re gonna output a table 
showing the mean and median steps per day

```{r show mean and median of steps per day in a table}
library(xtable)
xt <- xtable(data.frame(activity.daily))
print(xt, type="html")
```


## What is the average daily activity pattern?
For this question we have to go back to our dataset without NAs and group it 
again by 5-minute interval.

```{r create grouped dataset by 5-minute interval}
library(dplyr)
activity.grpd.int <- group_by(activity.wona, interval)
activity.int <- summarise(activity.grpd.int,
  avg_of_steps = sum(steps, na.rm = TRUE) / n())
```

Now we can make the plot with the already finished dataset which you can see here:

```{r plot the step interval graphic}
plot(activity.int$avg_of_steps ~ activity.int$interval, 
    type = "l",
    main = "Average amount of steps per 5-minute interval",
    xlab = "5-minute interval",
    ylab = 'Average amount of steps (overall)')
```


## Imputing missing values
1. How many NA values are there in the raw dataset?
```{r calculate number of NAs}
number_of_NA <- sum(!complete.cases(activity.raw))
```
There are `r number_of_NA`NA values in the dataset.

2. Which strategy for imputing the NAs will be used?
I will use the mean for the 5-minute-interval for imputing missing values.

3. Create a dataset based on the original dataset where the missing data is 
imputed.

```{r impute the missing data via a 5-minute-interval}
# make a LEFT OUTER JOIN to move the averages to the right side of the original
# dataframe
activity.merged <- merge(activity.raw, activity.int, by="interval", all.x=TRUE)

# now overwrite the NAs with the averages from the right side
activity.merged[is.na(activity.merged$steps), 2] <- activity.merged[is.na(activity.merged$steps), 4]
```

4. Based on 3 construct the grouped dataset for sum, mean and median of steps per day, plot a histogram and describe differences

```{r group imputed dataset}
# summarize the data
activity.imp <- group_by(activity.merged, date)
activity.imp.daily <- summarise(activity.imp,
  sum_of_steps = sum(steps, na.rm = TRUE),
  mean = mean(steps, na.rm = TRUE),
  median = median(steps, na.rm = TRUE))
```
Since we now got our working dataset we can start plotting

```{r histogram of imputed dataset}
hist(activity.imp.daily$sum_of_steps,
     main = "Histogram of imputed dataset of total steps",
     xlab = "sum of steps per day",)
```

For comparing the different means and medians of steps per day, I will create
a plot, where you can compare the different lines in the plot

```{r make compareable plot for mean and median of steps per day}
library(ggplot2)
ggplot()
```


## Are there differences in activity patterns between weekdays and weekends?
